---
title: Tasa de mortalidad en España por comunidades autónomas
author:
  - name: Adrián Carrasco Alcalá
    affil: 1, 2, \ddagger
  - name: Álvaro Nieva Valenzuela
    affil: 1, 3, \ddagger
  - name: Carlos David Llano Chicaiza
    affil: 1, 4, \ddagger
affiliation:
  - num: 1
    address: |
      Universitat de Valencia - Escuela Técnica Superior de Ingeniería (ETSE)
      Dirección: Avinguda de l'Universitat, S/N, 46100 Burjassot, Valencia
  - num: 2
    email: acal8@alumni.uv.es
  - num: 3
    email: alnieva@alumni.uv.es
  - num: 4
    email: cardalla@alumni.uv.es
# author citation list in chicago format
authorcitation: |
  Carrasco, A.; Nieva, A.; Llano C. D.
# firstnote to eighthnote
firstnote: |
  Estos autores contribuyeron por igual a este trabajo.
# document options
journal: data
type: article
status: submit
# front matter
simplesummary: |
  A Simple summary goes here.
abstract: |
  A single paragraph of about 200 words maximum. For research articles, 
  abstracts should give a pertinent overview of the work. We strongly encourage
  authors to use the following style of structured abstracts, but without 
  headings: 1) Background: Place the question addressed in a broad context and
  highlight the purpose of the study; 2) Methods: Describe briefly the main
  methods or treatments applied; 3) Results: Summarize the article's main 
  findings; and 4) Conclusion: Indicate the main conclusions or interpretations. 
  The abstract should be an objective representation of the article, it must not 
  contain results which are not presented and substantiated in the main text and 
  should not exaggerate the main conclusions.
keywords: |
  Tasa de mortalidad; Esperanza de vida; keyword 3.
# back matter
funding: |
  This research received no external funding.
dataavailability: |
  Todos los conjuntos de datos han sido obtenidos de la página del Instituto Nacional de Estadística (INE)
  (\url{https://www.ine.es/}). Son los siguientes:
  \url[Indicadores de renta media y mediana]{https://www.ine.es/jaxiT3/Tabla.htm?t=53689}
  \url[Renta media y mediana por comunidad autónoma]{https://www.ine.es/jaxiT3/Tabla.htm?t=68338}
conflictsofinterest: |
  The authors declare no conflict of interest. The founding sponsors had no role in the design of the study; 
  in the collection, analyses, or interpretation of data; in the writing of the 
  manuscript, an in the decision to publish the results
abbreviations:
  - short: MDPI
    long: Multidisciplinary Digital Publishing Institute
  - short: DOAJ
    long: Directory of open access journals
  - short: TLA
    long: Three letter acronym
  - short: LD 
    long: linear dichroism
bibliography: mybibfile.bib
appendix: appendix.tex
endnotes: false
output: 
  rticles::mdpi_article:
    extra_dependencies: longtable
---

# Introducción

The introduction should briefly place the study in a broad context and highlight
why it is important. It should define the purpose of the work and its
significance. The current state of the research field should be reviewed
carefully and key publications cited. Please highlight controversial and
diverging hypotheses when necessary. Finally, briefly mention the main aim of
the work and highlight the principal conclusions. As far as possible, please
keep the introduction comprehensible to scientists outside your particular
field of research. Citing a journal paper [@bertrand-krajewski_distribution_1998;
@leutnant_stormwater_2016]. And now citing a book reference @gujer_systems_2008.
Some MDPI journals use Chicago and others use APA, this template should choose
the correct citation format for you once you specify the journal in the YAML
header.

To use endnotes, change `endnotes: true` in the YAML header, then use
`\endnote{This is an endnote.}`.


# Materiales y Métodos

Materials and Methods should be described with sufficient details to allow 
others to replicate and build on published results. Please note that publication 
of your manuscript implicates that you must make all materials, data, computer 
code, and protocols associated with the publication available to readers. Please
disclose at the submission stage any restrictions on the availability of 
materials or information. New methods and protocols should be described in 
detail while well-established methods can be briefly described and appropriately
cited.

Research manuscripts reporting large datasets that are deposited in a publicly 
available database should specify where the data have been deposited and provide
the relevant accession numbers. If the accession numbers have not yet been
obtained at the time of submission, please state that they will be provided
during review. They must be provided prior to publication.

Interventionary studies involving animals or humans, and other studies require
ethical approval must list the authority that provided approval and the
corresponding ethical approval code.

# Resultados

This section may be divided by subheadings. It should provide a concise and
precise description of the experimental results, their interpretation as well
as the experimental conclusions that can be drawn.

## Subsection Heading Here

Subsection text here.

### Subsubsection Heading Here

Bulleted lists look like this:

* First bullet
* Second bullet
* Third bullet

Numbered lists can be added as follows:

1. First item
2. Second item
3. Third item

The text continues here.

## Figures, Tables and Schemes

All figures and tables should be cited in the main text as Figure \ref{fig:fig1}, 
\ref{tab:tab1}, etc. To get cross-reference to figure generated by R chunks 
include the `\\label{}` tag in the `fig.cap` attribute of the R chunk: 
`fig.cap = "Fancy Caption\\label{fig:plot}"`.

```{r fig1, echo=FALSE, fig.width=3, fig.cap="A figure added with a code chunk.\\label{fig:fig1}"}
x = rnorm(10)
y = rnorm(10)
plot(x, y)
```


When making tables using `kable`, it is suggested to use
the `format="latex"` and `tabl.envir="table"` arguments
to ensure table numbering and compatibility with the mdpi
document class.

```{r tab1, echo=FALSE}
knitr::kable(mtcars[1:5, 1:3], format = "latex", 
             booktabs = TRUE, 
             caption = "This is a table caption. Tables should be placed in the 
             main text near to the first time they are~cited.", 
             align = 'ccc', centering = FALSE,
             table.envir = "table", position = "H")
```



For a very wide table, landscape layouts are allowed.


\startlandscape

```{r tab2, echo=FALSE}
df <- data.frame(`Title 1` = c("Entry 1", "Entry 2"),
                 `Title 2` = c("Data", "Data"),
                 `Title 3` = c("Data", "Data"),
                 `Title 4` = c("This cell has some longer content that runs over
                               two lines",
                               "Data"))
knitr::kable(df, format = "latex", 
             booktabs = TRUE, 
             caption = "This is a very wide table", 
             align = 'ccc', centering = FALSE,
             table.envir = "table", position = "H")
```

\finishlandscape

## Formatting of Mathematical Components

This is an example of an equation:

$$
a = 1.
$$

If you want numbered equations use Latex and wrap in the equation environment:

\begin{equation}
a = 1,
\end{equation}

the text following an equation need not be a new paragraph. Please punctuate 
equations as regular text.

```{comment}
If the documentclass option "submit" is chosen, please insert a blank line before and after any math environment (equation and eqnarray environments). This ensures correct linenumbering. The blank line should be removed when the documentclass option is changed to "accept" because the text following an equation should not be a new paragraph.
```

This is the example 2 of equation:

\begin{adjustwidth}{-\extralength}{0cm}
\begin{equation}
a = b + c + d + e + f + g + h + i + j + k + l + m + n + o + p + q + r + s + t + 
u + v + w + x + y + z
\end{equation}
\end{adjustwidth}

Theorem-type environments (including propositions, lemmas, corollaries etc.) 
can be formatted as follows:

Example of a theorem:

::: {.Theorem latex=true}
Example text of a theorem
:::

The text continues here. Proofs must be formatted as follows:

Example of a proof:

::: {.proof latex="[Proof of Theorem1]"}
Text of the proof. Note that the phrase ``of Theorem 1'' is optional if it is 
clear which theorem is being referred to.
:::

The text continues here.

# Conclusiones

Authors should discuss the results and how they can be interpreted in 
perspective of previous studies and of the working hypotheses. The findings and 
their implications should be discussed in the broadest context possible. Future 
research directions may also be highlighted.
